---
title: "Introduction to the ltaer package"
author: "Shaun Khoo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `ltaer` package (pronounced 'later') is a convenient R wrapper for the various transportation-related APIs offered by the Land Transport Authority of Singapore (also known as the LTA Datamall). Currently, most of the APIs are geared towards app developers who use these real-time data for their mobile offerings. Used correctly and collected over a longer period of time, the data from the APIs can also serve as a rich dataset for people who are keen on analysing transportation trends in Singapore.

The key motivation behind this package is to make the data more accessible to a wider audience. While it is not difficult to call the API directly using the `httr` package on R, it does require a greater understanding of R and APIs that a casual user may not possess. An important part of open data is accessibility, and I hope that this package will constitute a small step towards that goal for Singapore.

This vignette will introduce you to the core functions within the `ltaer` package that interact directly with LTA's APIs. Another vignette will provide some examples of `plotSGMap`, which provides an interactive map visualisation of the data, and other examples of how this data can be visualised.

## 0. Working with LTA Datamall's APIs

####API Key

All API calls to the LTA Datamall, and thus all core functions in this package, require an API key. In layman terms, the API key is like a username/password with which you access the API. It helps the API developers to identify who is accessing the API, and to restrict access if they wish. As such, it is really important that you keep your API key **private** (don't share it with anyone else or post it online!). 

As such, in order to use this package, please sign up for an API key [here](https://www.mytransport.sg/content/mytransport/home/dataMall/request-for-api.html). The process is quick, easy and straightforward. 

You will notice that in the examples provided below, my API key is saved as `Sys.getenv('LTA_DATAMALL_KEY')`. This is one way to keep the API key private, and you can find out more about how to do this [here](https://daattali.gitbooks.io/stat545-ubc-github-io/bit003_api-key-env-var.html).

####API Documentation

Finally, please read this vignette alongside LTA's documentation for their APIs, which you can find [here](https://www.mytransport.sg/content/dam/datamall/datasets/LTA_DataMall_API_User_Guide.pdf). This will help you understand the outputs you are receiving from the APIs you are calling. By default, all functions in `ltaer` return all the data that the API returns, so you do not have to worry about missing out on any data from using this package.

####Installing this package

This package is not available on CRAN (this is where you typically download R packages from when you run `install.package('pkgname')`). Due to a lack of time on my end, it is unlikely that `ltaer` will ever be made available on CRAN either. The only way you can get `ltaer` working on your RStudio is to use the following code:

```{r, eval=FALSE}
install.packages('devtools')
devtools::install_github('shaunkhoo/ltaer')
```

This will download the `ltaer` package directly from my Github repository and into your RStudio. Thereafter, you simply load the package in the same way as any other package, by entering `library(ltaer)`.

## 1. Bus Arrival Timings

Probably the most well-utilised API of the lot, the Bus Arrival API enables users to retrieve bus arrival timings for any bus stop in Singapore. There are some specific things to mention here about the API (you can also find this in the documentation):

* For each bus service serving that bus stop, the bus arrival details (timing, coordinates, type, load etc) are reported for the next three buses. If there are only one or two 'next buses' left (usually around 11pm-1am), then the bus arrival details for those 'next bus' slots will be NA. 
* All bus services where the last bus has passed the chosen bus stop will no longer show up on the Bus Arrivals API for that bus stop. This usually happens around 12am-1am, except for NightRider services which end at 3am-4am.
* If there are no bus services at all, then the API will return an empty JSON file.

There are some important differences between this package's wrapper function, `getBusTimings`, and the API itself:

* `getBusTimings` automatically calls all the API for all bus services for a given bus stop code. In the API, you can specify which service you want to check for each bus stop. I've dropped this option as R users can simply filter for the bus service number.
* `getBusTimings` conveniently allows users to call the API for multiple bus stops. This is done by simply repeating the API call for all the bus stops specified. You simply need to pass in a vector of bus codes and `getBusTimings` will return you a dataframe containing all the data returned. The first column, `BusStopCode`, will help you distinguish between the various bus stops.
* `getBusTimings` requires you to pass the bus stop code as a character string, and not as a numeric value. This is to avoid potential issues when the bus stop code has leading zeros, which R truncates automatically for numeric values.

I provide an example below of the output you can expect from `getBusTimings`.

```{r}
library(ltaer)

# The following code is commented out as it may not work all the time. The output has been saved in this package as 'bus_arrivals' for your use.
# bus_arrivals <- getBusTimings(c('07531', '09047', '11239', '14151', '66451', '44289', '50021', '54269', '03019', '09022', '41029'), api_key = Sys.getenv('LTA_DATAMALL_KEY'))

# Note that I've transposed the dataframe (the columns are the rows and vice versa) for clarity of presentation.
knitr::kable(t(bus_arrivals[1:3,]))
```

## 2. Bus Services

This API returns service information for all buses currently in operation in Singapore. The data available here is similar to those you find on the information boards at each bus stop.

`getBusServices` is the wrapper function for this API, and takes in the API key as its only argument.

```{r warning=FALSE}
bus_services <- getBusServices(Sys.getenv('LTA_DATAMALL_KEY'))

# Note that I've transposed the dataframe (the columns are the rows and vice versa) for clarity of presentation.
knitr::kable(t(bus_services[1:9,]))
```

## 3. Bus Routes

This API returns route information for all services currently in operation in Singapore. The data here is quite useful as they provide each bus stop code and stop sequence for each bus service, so you can plot out the full route for each bus service. While this is already quite useful, it would have been even better if route polylines were provided as well.

`getBusRoutes` is the wrapper function for this API, and takes in the API key as its only argument.

```{r}
# The following code is commented out as it takes a long time to retrieve all the routes. The output has been saved in this package as 'bus_routes' for your use.
# bus_routes <- getBusRoutes(Sys.getenv('LTA_DATAMALL_KEY'))

# Note that I've transposed the dataframe (the columns are the rows and vice versa) for clarity of presentation.
#knitr::kable(bus_routes[1:9,])
```

## 4. Bus Stops

This API returns information for all bus stops currently being serviced by buses in Singapore, such as bus stop codes, coordinates and road names.

`getBusStops` is the wrapper function for this API, and takes in the API key as its only argument.

```{r message=FALSE}
bus_stops <- getBusStops(Sys.getenv('LTA_DATAMALL_KEY'))
knitr::kable(head(bus_stops))
```

## 5. Passenger Volume (Buses)

There are two 




## 6. Passenger Volume (Trains)



## 7. Taxi Availability

This API returns the coordinates for all available taxis that are available for hire in Singapore (excludes hired and 'busy' taxis). 

`getTaxiAvail` is the wrapper function for this API, and takes in the API key as its only argument.

```{r message=FALSE}
taxi_avail <- getTaxiAvail(Sys.getenv('LTA_DATAMALL_KEY'))
knitr::kable(head(taxi_avail))
```

## 8. Carpark Availability

This API returns the coordinates for all available taxis that are available for hire in Singapore (excludes hired and 'busy' taxis). 

`getTaxiAvail` is the wrapper function for this API, and takes in the API key as its only argument.

```{r message=FALSE}
taxi_avail <- getTaxiAvail(Sys.getenv('LTA_DATAMALL_KEY'))
knitr::kable(head(taxi_avail))
```







