---
title: "ltaer: Visualisations"
author: "Shaun Khoo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 0. Visualising Geospatial Data

As mentioned in the introduction, the main purpose of the `ltaer` package is to improve accessibility of Singapore's transportation data to the wider public. Visualisations are a crucial part of this because it communicates key observations and trends from the data in an intuitive and comprehensible manner. 

Visualising geospatial data is generally more challenging that running visualisations on the `ggplot2` package, primarily because there are more things one has to deal with. The good thing is that there are many R packages that greatly simplify the amount we need to learn and do to create effective visualisations.

Hence, this vignette aims to provide some suggestions and tips in visualising the data obtained from the various APIs. I begin first with introducing `plotSGMap`, a Leaflet wrapper provided in this package that enables you to plot your data with minimal coding.

## 1. Using plotSGMap

`plotSGMap` is a convenient Leaflet wrapper which allows users to quickly plot geospatial data on an interactive map for data exploration. A picture speaks a thousand words, so...

``` {r}
library(ltaer)
plotSGMap(bus_stops, popup = 'Description', cluster = FALSE, colour = 'green', alpha = 1, size = 1)
```

The key contributions here are twofold: users do not need any knowledge of `leaflet`, and `ggplot2` grammar is used here for familiarity (`alpha` and `size` are not `leaflet` arguments). The motivation here is to enable users to have a quick look at the data from a spatial perspective, and thus to quickly get to grips with the data they have received from the APIs.

Here is an explanation of the various arguments in the function:

* `dataframe`: The dataframe you want to plot on the Singapore map. It must contain one column for latitude, named either as 'lat' or 'latitude', and one column for longitude, named either as 'lng' or 'longitude' - neither are case-sensitive. 
* `popup`: What appears when you click on each point on the Singapore map. It must be a character string and name another column in the `dataframe` that you want to display in the pop up.
* `cluster`: Whether you want to display each point individually or cluster them automatically based on regions. By default, this is set to `FALSE`.
* `colour`: The colour of each point being plotted. You can either state some simple colours directly (like 'red' or 'blue') or use a hex code - the default here is 'red'.
* `alpha`: How opaque each point is. You can set it to 1 for fully opaque, and 0 for fully transparent - the default here is 0.5.
* `size`: How large each point is. This is set to 5 by default, although for more cluttered maps you may want a smaller value like 1.








