---
title: "ltaer: Visualisations"
author: "Shaun Khoo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 0. Visualising Geospatial Data

As mentioned in the introduction, the main purpose of the `ltaer` package is to improve accessibility of Singapore's transportation data to the wider public. Visualisations are a crucial part of this because it communicates key observations and trends from the data in an intuitive and comprehensible manner. 

Visualising geospatial data is generally more challenging that running visualisations on the `ggplot2` package, primarily because there are more things one has to deal with. The good thing is that there are many R packages that greatly simplify the amount we need to learn and do to create effective visualisations.

Hence, this vignette aims to provide some suggestions and tips in visualising the data obtained from the various APIs. I begin first with introducing `plotSGMap`, a Leaflet wrapper provided in this package that enables you to plot your data with minimal coding.

## 1. Using plotSGMap

A picture speaks a thousand words, so:

``` {r}
library(ltaer)
plotSGMap(bus_stops, popup = 'Description', cluster = FALSE, color = 'red', alpha = 1, size = 1)
```



Leaflet is one of the






